Patient-Specific Neural Field Model

### 2010-07-19 ltx/data_paper/DataPaper.tex
began to add a model derivation to the paper and updated figures

\section{Introduction}
\begin{itemize}
	\item Understanding of macroscopic neurodynamics
	\item Neural field model
	\item Improvements in electrode technology with great spatiotemporal resolution
\end{itemize}

\section{Method}

\subsection{Data Collection and Pre-processing}
\begin{itemize}
	\item Show medical imaging
	\item Show LFP Data (background, seizure, post-seizure)
	\item Show spectral properties of LFPs
\end{itemize}

@@

\subsection{Spatio-temporal Analysis}
The first 5 channels of the LFP time series from the seizure is shown in Figure~\ref{fig:TimeSeries}. For the analysis, the time series was segmented into three periods; the pre-seizure period, seizure period and post-seizure period. The seizure onset is marked by the red dotted line and the seizure end is marked by the blue dashed line. An example to illustrate the spatial aspects of the observed neural field is shown in Figure~\ref{fig:FieldObserations}. Each subplot in the figure shows the observaed field at a sample of the pre-processed data, with each plot separated in time by 2.5~ms. The figure shows spatially correlated activity.

@@

@@

To confirm the applicability of the model spatio-temporal analysis were performed. The spectral properties of the recorded time series for each of the time segments are shown in Figure~\ref{fig:TemporalFreqObservation}.

@@

\subsection{Frequency Analysis}
\begin{itemize}
	\item Create 2D FT of observations to estimate bandwidth
\end{itemize}

@@

@@

\bibliographystyle{plain}
\bibliography{}
\end{document}

>>>>

\section{Introduction}
\begin{itemize}
	\item Understanding of macroscopic neurodynamics
	\item Neural field model
	\item Improvements in electrode technology with great spatiotemporal resolution
\end{itemize}
The aim of this paper is to great a patient-specific neural field model, by estimating parameters of a general mathematical model using data acquired from an epilepsy surgical candidate in a clinical setting. 
\section{Method}

\subsection{Neural Field Model}
The neural field equations used as the parametric form of the model are based on the influential papers from Wilson and Cowan~\cite{Wilson1973}, and Amari~\cite{Amari1977}. This class of model describes a continuous cortical sheet or surface, relating mean firing rates of pre and post-synaptic neural populations. The neural field, $v\left( {\mathbf{r},t} \right)$, at position $\mathbf{r}$ and time $t$, is the aggregated post-synaptic potentials described by the temporal convolution 
\begin{equation}
	\label{SpikesToPotential} v\left( {\mathbf{r},t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'}, 
\end{equation}   
where $h(\cdot)$ is the synaptic response kernel and $g(\cdot)$ describes the mean firing rate. The firing rate is described by spatial convolution
\begin{equation}
	\label{RateBasedInteractions} g\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)\textrm{d}\mathbf{r}'}, 
\end{equation}
where $f(\cdot)$ relates mean post-synaptic potentials to mean firing rates, $w(\cdot)$ is the spatial connectivity kernel and $\Omega$ is the spatial domain representing the cortical sheet or surface. The firing rate function is the sigmoidal activation function described by
\begin{equation}
	\label{ActivationFunction} f\left( v\left( \mathbf{r}', t \right) \right) = \frac{1}{1 + \exp \left( \varsigma \left( v_0 - v\left(\mathbf{r}',t\right) \right) \right)}, 
\end{equation}
where $v_0$ and $\varsigma$ describe the firing threshold the slope of the sigmoid respectively. By substituting equation~\ref{RateBasedInteractions} into equation~\ref{SpikesToPotential} we get the spatiotemporal model 
\begin{equation}
	\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) =
	\int_{-\infty}^t 
	h\left(t - t'\right) \int_\Omega
	w\left(\mathbf{r},\mathbf{r}'\right) 
	f\left( v\left( \mathbf{r}',t' \right)\right)
	\textrm{d}\mathbf{r}'dt'.
\end{equation}
To arrive at the general intergral-differential equation form of the model, we shall express the synaptic response kernel as a Green's function 
\begin{equation}
	\label{GreensFuncDef} Dh\left( t \right) = \delta \left( t \right), 
\end{equation}
where $D$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{equation}
	\label{FinalFormContinuous} 
	Dv\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)\textrm{d}\mathbf{r}'}. 
\end{equation}
\subsection{Data Collection and Pre-processing}
\begin{itemize}
	\item Show medical imaging
	\item Show LFP Data (background, seizure, post-seizure)
	\item Show spectral properties of LFPs
\end{itemize}

@@

\subsection{Spatiotemporal Data Analysis}
The first 5 channels of the LFP time series from the seizure is shown in Figure~\ref{fig:TimeSeries}. For the analysis, the time series was segmented into three periods; the pre-seizure period, seizure period and post-seizure period. The seizure onset is marked by the red dotted line and the seizure end is marked by the blue dashed line.

@@

An example to illustrate the spatial aspects of the observed neural field is shown in Figure~\ref{fig:FieldObserations}. Each subplot in the figure shows the observed field at a sample of the pre-processed data, with each plot separated in time by 2.5~ms. The figure shows spatially correlated activity.

@@

The spectral properties of the recorded time series for each of the time segments are shown in Figure~\ref{fig:TemporalFreqObservation}. Each plot shows an averaged (across channels) spectrum the data. The pre-seizure period shows a decay in power at a typical $1/f^2$ rate. The seizure period has higher power at low and high ends of the spectrum. In addition, there is a peak in the power of the LFPs at ~51~Hz during the seizure. This peak would be difficult to observe in European and Asian countries due to the 50~Hz mains power frequency. The post-seizure period shows a large increase in lower frequency power...

@@

Figure~\ref{fig:SpatialFreqObservation} shows the results of spatial frequency analysis of the observed neural field. Throughout this paper the spatial frequency shall be denoted by $\nu$. Each subplot is an average (across time) spatial frequency during either the pre-seizure, seizure or post-seizure periods. The spatial frequency of the neural field is governed by disturbances (inputs from outside the field) and the underlying connectivity structure. The symmetry in the plots provides some evidence for homogeneous and isotropic connectivity.

@@

Figure~\ref{fig:DiagSpatialFreqObservation} shows diagonal cross-sections of the spatial frequency plots. From this figure we can see that the spatial cut-off frequency (-3~dB point) is $\nu_c \approx 0.26$~Hz. The profile of the diagonal cross-section indicates that the field is adequately sampled from the micro-electrode array, where the power drops by ~25~dB at the maximum observable spatial frequency (according to Shannon's sampling theorem) indicating spatial aliasing will have an insignificant effect. The spatial cut-off frequency provides information on the complexity of the model required to represent the neural field. Considering the field is band-limited, we can represent the continuous field by a finite set of continuous basis functions. According the Shannon's sampling theorem, the minimum spacing between basis functions, $\Delta_{\phi}$, to describe the field is
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_{\phi} \leq \frac{1}{2\rho_{\phi}\boldsymbol{\nu}_{c}}
\end{equation}
where $\rho_{\phi} \in \mathbb{R} \ge 1$ is an oversampling parameter to determine the basis function separation.

@@

\bibliographystyle{plain}
\bibliography{BrainIDE}
\end{document}
